---
title: "Wrangling categorical variables in R"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Explore the data

```{r warning = FALSE, message = FALSE}
library(tidyverse)
pokemon <- read_csv("data/pokemon.csv")
```

- How many different types of pokemon are there?
- How many body styles do they have?

_WRITE YOUR ANSWERS HERE_

_Hint:_ try using functions like `str()` or `summary()` to get an overview of the data. For a more comprehensive overview, try the `skimr` package.

```{r eval = FALSE}
# install.packages("skimr")
library(skimr)
skim(pokemon)
```

## Let's get graphing

We can start by changing the `ggplot2` graphing defaults. If you set it once now, you don't have to add `+ theme_minimal()` or `+ scale_fill_viridis_c()` to your plots later on. 

_Note:_ these settings do not always work for `ggpubr` graphs.

```{r}
# change the default theme
theme_set(theme_minimal())

# change the default colors
options(ggplot2.continuous.colour = "viridis",
        ggplot2.continuous.fill = "viridis")
```

How do you change the code below to match the [graph](https://ecodatasci-tlv.github.io/factors/#27) shown in the slides?

```{r}
# YOUR CODE HERE

pokemon %>% 
    # hint: add some mutates here to change the factor order
    ggplot(aes(type_1)) +
    geom_bar() + coord_flip()
```

How do you recreate the [graph](https://ecodatasci-tlv.github.io/factors/#31) shown in the second plotting exercise?

```{r}
# YOUR CODE HERE
```

## Discuss

- How is `fct_infreq()` different from `fct_reorder()`? How are they similar?

_WRITE YOUR ANSWER HERE_



